Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-03-05 latex/MPP21/paper.tex
drafted conclusion section

\begin{figure}
\subfloat[One thread]{
\centering
\renewcommand{\datafile}{data/NN_seq.dat}
\input{plots/NN_reltime}
\label{fig:NNreltime_seq}}
\\
\subfloat[Twelve threads]{
\centering
\renewcommand{\datafile}{data/NNtime_12t.dat}
%\input{plots/NN_reltime}
\label{fig:NNreltime_par}}
\caption{Network training time relative to using classical matrix multiplication}
\label{fig:NNreltime}
\end{figure}

@@

\begin{enumerate}
	\item promising results for synthetic network, goal is to target costly networks bottlenecked by matmul (\GB{\cite{NPOV15} uses a network including layers with dimensions 25K x 4K and 4K x 4K})
	\item can tune for matmul dimensions (not just square)
	\item GPU implementation (cite Jianyu Huang's work)
	\item mixed precision fast matmul (APA or otherwise)
\end{enumerate}

>>>>

\begin{figure}
\subfloat[One thread]{
\centering
\renewcommand{\datafile}{data/NN_seq.dat}
\input{plots/NN_reltime}
\label{fig:NNreltime_seq}}
\\
\subfloat[Six threads]{
\centering
\renewcommand{\datafile}{data/NN_par_6.dat}
%\input{plots/NN_reltime}
\label{fig:NNreltime_par_6}}
\\
\subfloat[Twelve threads]{
\centering
\renewcommand{\datafile}{data/NN_par_12.dat}
%\input{plots/NN_reltime}
\label{fig:NNreltime_par_12}}
\caption{Network training time relative to using classical matrix multiplication}
\label{fig:NNreltime}
\end{figure}

@@

Arbitrary Precision Approximating algorithms offer significant practical performance improvements over exact fast matrix multiplication algorithms.
They outperform classical matrix multiplication at smaller matrix dimensions and yield greater benefits as matrix dimensions increase.
APA algorithms sacrifice accuracy, achieving error of only a fractional root of the working precision, but they can be practical for applications that do not demand high accuracy.
The results presented here show that neural network training, particularly for Multi-Layer Perceptron networks, is robust to such matrix multiplication error, and APA algorithms are a useful tool for accelerating the costly training computations.
We observe that for APA algorithms offering a theoretical speedup of up to 39\%, we can achieve up to 28\% and 21\% improvements for single-threaded and multi-threaded implementations of individual matrix multiplications, respectively, which translate to up to \GB{W\%} speedup in overall training time on a multicore CPU.

While the accuracy and performance results are promising for MLP networks and the MNIST dataset, we believe that APA algorithms can also be effective in state-of-the-art networks and more difficult machine learning problems.
For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers, including a layer with dimensions $25088\times 4096$, where APA algorithms can offer significant performance improvement.

As illustrated by Benson and Ballard \cite{BB15}, the highest performing fast algorithms for rectangular matrix multiplications often have dimensions that match the aspect ratio of the problem.
We consider in this paper only square matrix multiplications, and we see that square (or nearly square) dimensions of $\dims{4,4,4}$ and $\dims{4,4,2}$ are fastest.
Algorithms with more skewed aspect ratios will likely perform better for problems with matching skewed matrix multiplication dimensions.
We note that an algorithm for dimensions $\dims{m,n,k}$ can be translated into an algorithm for $\dims{n,m,k}$ and any other reordering of the dimensions \cite{BB15}.
For very large matrices, it may also be beneficial to consider more than one recursive step of the same algorithm or to use a combination of two or three different algorithms across recursive steps (i.e., uniform, non-stationary algorithms \cite{BBDLS16}).

Finally, we would like to extend the code generation techniques to other hardware platforms, including GPUs.
Because GPUs can offer higher efficiency for matrix multiplications, we wish to demonstrate the potential of APA algorithms in that environment.
Techniques used to accelerate Strassen's algorithm on the GPU \cite{HYvdG20,LAES13} can be generalized to other fast algorithms as done in work for multicore CPUs \cite{BB15,HRMvdG17}.
